Exploring Multi-Scale Spatiotemporal Twitter User Mobility Patterns with a Visual-Analytics Approach

### 2016-08-14 MovePattern_IJGI/IJGI/Twitter_MovementPatterns_VisualAnalytics.tex
new
xx
new drive

%=================================================================
% Full title of the paper (Capitalized)
\Title{Exploring multi-scale spatiotemporal Twitter user mobility patterns with a scalable visual-analytics framework}

% Authors, for the paper (add full first names)
\Author{Junjun Yin $^{1}$ and Zhenhong Du $^{2}$*}
% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

@@

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{Understanding human mobility patterns is of great importance for urban planning, traffic management, and even marketing campaign.
However, the capability of capturing detailed human movements with fine-grained spatial and temporal granularity is still limited.
In this study, we have extracted high-resolution mobility data from a collection of 1.3 billion geo-located Twitter messages to study multi-scale spatiotemporal Twitter user mobility patterns in the United States during the year 2014. 
Regarding the concerns of infringement on individual privacy, such as the mobile phone call data with restricted access, the dataset is collected from publicly accessible Twitter data streams.
In this study, we have developed a scalable visual-analytics framework to deliver efficiency and scalability in filtering large volume of geo-located tweets, modeling and extracting Twitter user movements, generating space-time user trajectories, and summarizing multi-scale spatiotemporal user mobility patterns.
We have performed a set of statistical analysis to understand Twitter user mobility patterns across multi-level spatial scales and temporal granularity.
In particular, the Twitter user mobility patterns measured by the displacements and radius of gyrations of individuals have revealed multi-scale or multi-modal Twitter user mobility patterns.
By further studying such mobility patterns in different temporal ranges, we have identified both consistency and seasonal fluctuations regarding the distance decay effects in the corresponding mobility patterns.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Understanding human mobility patterns is of great importance for a broad range of applications from urban planning~\cite{zheng2008understanding}, traffic management~\cite{jiang2009characterizing}, and even the spatial spread of epidemic diseases~\cite{belik2011natural}.
Earlier research efforts relied on low resolution mobility data to understand human mobility patterns, such as using census records to understand human migration patterns~\cite{greenwood1985human}, or delivering questionnaires and asking volunteers to report the track of bank notes to infer human travel patterns~\cite{brockmann2006scaling}.
However, the lack of detailed human movements with fine-grained spatial and temporal granularity limits the findings to capture mobility patterns of individuals~\cite{gonzalez2008understanding,Jurdak2015}.
In addition to the mobility data collected by GPS trackers~\cite{zheng2008understanding, rhee2011levy} and mobile phone call records~\cite{gonzalez2008understanding,sevtsuk2010does,kung2014exploring}, 
emerging as a new source for mobility data, today's pervasive Location Based Social Media (LBSM) platforms (e.g., Twitter and Foursquare) offer continuous spatial Big Data streams with massive amount of detailed and frequently updated user digital footprints~\cite{thatcher2014living} in the form of real-world trails and footprints.
One significant advantage of LBSM data streams is the large spatial coverage, for example, researchers have used geo-located Twitter data for studying global mobility patterns~\cite{hawelka2014geo}, which is otherwise impossible by using other mobility datasets (e.g., GPS traces and mobile phone call records). 
In addition, the publicly available LBSM data streams offer unique opportunities for conducting replicable scientific findings regarding the concerns of infringement on individual privacy, such as using mobile phone call records~\cite{giannotti2008mobility,crampton2014collect,Jurdak2015}.

Many studies have adopted the LBSM data streams to study human mobility patterns. For example, they modeled and extracted trajectories of individuals and performed statistical analysis focusing on the distance decay effects~\cite{gonzalez2008understanding} in the collective user movements to reveal different travel modes~\cite{Jurdak2015}, travel demands~\cite{wu2014intra,hasan2013understanding}, and the impact of social connections~\cite{cho2011friendship}.
These studies have provided strong support for using LBSM data as proxies for studying mobility patterns of individuals and valuable insights into human mobility dynamics. 
However, in these studies, the measurements of distances are either fixed in a certain time period or a specific region, where they do not consider the variations of movements in different spatial scales and temporal ranges.
To be more specific, these studies lack examinations of the detailed movements across different geographical scales and temporal granularity, for instances, whether there are temporal (e.g., monthly or seasonal) changes within the movements, and how the mobility patterns vary across different spatial scales (e.g., intra- or inter county, city or country level). 
These insights are critical to advance our understandings of the collective mobility patterns for a variety of applications, such as examining the mobility patterns across different cities~\cite{noulas2012tale}, the spread patterns of disease~\cite{balcan2009multiscale, tamerius2011global} and touristic activities~\cite{hawelka2014geo}.
While the high resolution spatiotemporal records from LBSM present unique research opportunities in this direction, the inherited large data volume poses significant data intensive challenges for developing a multi-scale spatiotemporal analysis framework~\cite{tsou2015} to deal with the complexities in filtering movements of individuals, modeling and aggregating user trajectories at multiple spatial and temporal scales.
In addition, to enable efficient analysis and visualization of the mobility patterns at different spatiotemporal scales, it is essential for such a framework to provide scalability for addressing the data intensive challenges~\cite{cao2014scalable}.

In this paper, we have explored and studied the Twitter user mobility patterns across multi-level spatial scales and temporal granularity in the United States during the year 2014. 
The mobility data is extracted from1.3 billion geo-located Twitter messages (i.e., tweets) from 1$^{st}$ January to 31$^{st}$ December, 2014 in the United States with over 6 million Twitter users and over 1 TB in file size. 
To address the data-intensive challenges embedded in this dataset, we have developed a scalable visual-analytics framework tailored to accommodate large volume of geo-located tweets for studying multi-level spatiotemporal Twitter user mobility patterns.
This framework is implemented based on high-performance distributed computing environment using Apache Hadoop\footnote{http://hadoop.apache.org/}, which is an open source software framework to enable distributed processing of large datasets across computing clusters.
%We have implemented a set of efficient MapReduce~\cite{dean2008mapreduce} algorithms to deliver scalability in filtering large volume of geo-located tweets, modeling and extracting Twitter user movements, generating space-time user trajectories, and summarizing multi-level spatiotemporal user mobility patterns.
With this framework, we have performed a set of statistical analysis to understand the spatiotemporal Twitter user mobility patterns. 
We have modeled the frequency of Twitter users visiting different locations to study the collective user visiting behaviors, where we have identified temporal similarities in the distributions. 
In particular, the Twitter user mobility patterns measured by the displacements and radius of gyrations of individuals~\cite{gonzalez2008understanding} have revealed different groups of Twitter users with multi-scale or multi-modal mobility patterns and multiple travel modes~\cite{Jurdak2015}.
By further studying such mobility patterns in different temporal ranges, we have identified both consistency and seasonal fluctuations regarding the distance decay effects in the corresponding mobility patterns.

The remainder of this paper is organized as follows.
Section 2 describes the related work in the context of studying mobility patterns using LBSM data, in particular, geo-located Twitter data.
We focus on research challenges in visual-analytics methods to enable multi-scale
spatiotemporal analysis with massive movement datasets, including data management, multi-level spatiotemporal user trajectory modeling and visualization.
Section 3 details the processes for extracting, aggregating and summarizing multi-level spatiotemporal Twitter user mobility patterns.
Section 4 presents the case study of performing visual-analytics for seeking Twitter mobility patterns in the United States of year 2014. Section 5 concludes the paper.

\section{Mobility patterns in Location Based Social Media data}
\subsection{Geo-located Twitter data for studying large-scale user movements}
To understand detailed human mobility patterns of individuals, the capability of capturing human movements with fine-grained spatial and temporal granularity is critical.
However, the low-resolution mobility data collected from census records~\cite{greenwood1985human} is estimated and aggregated at census tract level, while tracking of bank notes~\cite{brockmann2006scaling} is at zip code level and do not necessarily reflecting movements of the same individuals~\cite{gonzalez2008understanding}.
In terms of collecting detailed mobility data of individuals, using GPS trackers tends to produce, to date, the most accurate records of individuals' movements regarding the accuracy of recorded user locations and update frequency~\cite{zheng2008understanding}. However, the data is often limited in spatial scale (e.g. within a specific city or region) with a small group of people, for example, 182 and 226 volunteers participated in collecting such mobility data in~\cite{zheng2010geolife} and~\cite{rhee2011levy} respectively. Other than tracking people directly, the vehicle-based GPS data is often tied to a specific vehicle (e.g. taxi), which may only be accessible for a certain group of people~\cite{kung2014exploring}.

Another approach from the literatures for studying human mobility is using mobile phone call data in the form of Call Detail Records (CDR), where the locations of mobile users are estimated by cell tower triangulation with accuracy in the order of kilometers~\cite{gonzalez2008understanding,sevtsuk2010does,kung2014exploring}. 
Such a dataset can cover relatively large spatial scale ~\cite{becker2013human,sobolevsky2013delineating} (e.g., national level) and a large portion of the population in the study region~\cite{kung2014exploring}. However, due to the concerns of infringement on individual privacy, the mobile phone call data is not publicly accessible, which is not ideal  for conducting replicable scientific findings, such as validating or extending the existing discoveries.

In this connection, it becomes increasingly popular for researchers to exploit the publicly accessible mobility data captured from today's pervasive Location Based Social Media (LBSM) platforms (e.g., Foursquare and Twitter). 
LBSM enables users to attach their current location as a geo-tag to the message they will post, which is derived from either the GPS or Wi-Fi positioning with a high position resolution down to 10 meters~\cite{Jurdak2015}.
A Big Data scenario emerges when millions social media users constantly posting messages.
However, there are some limitations and complexities in directly using the LBSM data for study human mobility patterns. 
For example, comparing to GPS traces, the update frequency of an individual's location varies depending on when a user is posting a new geo-located message or check-in at a new place. It has been criticized for the lack of representativeness of the population as not all people use social media or send geo-located messages~\cite{kung2014exploring}. Another research challenge is to identify the users, as a social media account is not equal to a real person in the physical world~\cite{tsou2015}.
Although many studies started to look into the demographic information of LBSM data, in particular Twitter data~\cite{mitchell2013geography,longley2015geotemporal}, these issues certainly require us to pose more strict criteria in filtering and extracting individual movements.

In this study, geo-located Twitter data is chosen as source for studying detailed mobility patterns. Compared to other LBSM platforms, Twitter is one of the most popular platforms and is been actively used in many countries. It provides a publicly accessible streaming API\footnote{https://dev.twitter.com/streaming/overview} for easy access to its data, in fact, many other LBSM data can be collected from the data streams, such as Foursquare check-in data~\cite{cranshaw2012livehoods, hasan2013understanding}.
In addition, it presents some unique advantages that make it suitable for studying human mobility patterns.
For example, the high-resolution location information enables to identify multiple travel modes in user mobility patterns~\cite{Jurdak2015}; the large spatial coverage enables to study global mobility patterns~\cite{hawelka2014geo}, which is almost impossible for other mobility datasets.
More importantly, by continuously monitoring the geo-located Twitter data streams with large volumes of detailed and frequently updated spatiotemporal  records of Twitter users, it offers a great deal of potential for studying mobility patterns of large groups of individuals at different spatial scales (e.g., movements across cities, states or even countries) and temporal gratuity (e.g., weekly, monthly, and seasonal movements).

\subsection{Data-intensive challenges for multi-scale geo-visual analytics}
Mobility data is essentially a collection of spatiotemporal records of people moving from one location to another in the geographic space.
To study mobility patterns of individuals, a space-time trajectory of each individual should be modeled and constructed to quantify the collective movements over space and time.
Based on the extracted space-time trajectories, aforementioned studies are able to perform analysis, such as the measurements of displacements and radius of gyrations of individuals, to study the mobility dynamics.
At the same time, space-time trajectory is one of the core concepts in H{\"a}gerstrand's time geography to understand the embedded spatiotemporal dynamics~\cite{hagerstrand1985time}, which has provided useful insights to explore movements across different geographical scales and temporal granularity.
For example, a geo-visualization approach was used to study human activity patterns, where user trajectories are mapped in a 3D space ordered by timestamps in the third dimension~\cite{kwan2004geovisualization}. While such an approach enables visualization of detailed trajectories, its capability is limited in dealing with large-volume movement datasets~\cite{andrienko2007designing}.
Instead of directly visualizing individual user trajectories, a space-time cube approach was proposed in analyzing and visualizing the collective trajectories by providing flexibilities in setting up both spatial and temporal ranges, and therefore to study the mobility patterns across different spatial units (e.g. countries, states, and cities, etc.) and identify the changes over space and time~\cite{maceachren2001research, maceachren2004maps}.
In this connection, visual-analytics methods are proposed to help better convey the findings in terms of analyzing and visualizing multi-level spatiotemporal mobility patterns~\cite{andrienko2007designing,andrienko2007visual}.
Visual-analytics methods focus on the synergy of computational and analytical methods to reduce the visual clutter, where aggregation methods are suggested to perform grouping/dividing individual's moving trajectories at different spatial and temporal granularity, e.g., utilizing the space-time cube approach~\cite{andrienko2007designing}.

Employing visual-analytics methods dealing with massive movement datasets is not only beneficial for optimizing visualizations but also provides a great deal of flexibilities for performing statistical analysis in seeking mobility patterns with different level of spatiotemporal details. 
By conveying the findings from statistical analysis, it provides another visualization form regarding the corresponding mobility patterns, in a sense, to avoid the problems of visual clutters.
However, in the context of studying mobility patterns using large volume of geo-located Twitter data, the inherited large data volume poses significant data intensive challenges for these mentioned visual-analytics methods to scale with both the data volume and the computational requirements (e.g., movement extraction and trajectory modeling)~\cite{cao2014scalable}.
For example, in our study, 1.3 billion geo-located tweets were collected with over 1 TB in file size.
To construct a space-time trajectory of an individual, it is necessary to go through the massive dataset to sort and update the trajectory whenever a new location is found. 
Such a task is already computationally demanding, let along breaking the trajectories to construct space-time cube with multiple spatial scale and temporal ranges. 
In particular, developing a multi-scale spatiotemporal analysis framework is identified as one of the research challenges for dealing with social media Big Data~\cite{tsou2015}. 
To address the data-intensive challenges, there is a need to develop a scalable visual-analytics framework tailored to accommodate large volume of geo-located tweets for studying multi-level spatiotemporal Twitter user mobility patterns.

%\subsection{Accommodating LBSM Big Data with Hadoop}
%Apache Hadoop is an open source software framework designed to facilitate data intensive computing on large commodity clusters.
%It combines a distributed file system, namely Hadoop Distributed File System(HDFS)~\cite{shvachko2010hadoop} with MapReduce programming paradigm, which can be applied to wide range of data-intensive problems.
%MapReduce is a programming model and an associated software framework designed to process massive data in a distributed fashion~\cite{dean2008mapreduce}.
%MapReduce breaks the entire computation into small tasks and schedule them among different computing nodes.
%MapReduce consists of two main stages: map and reduce.
%In the map stage, input data is converted into series of intermediate $\langle key, value\rangle$ pairs.
%Customized computation will be performed in the reduce stage based on the same intermediate key.
%This framework provides parallelization since both map and reduce tasks are considered independent and can be done in parallel. More importantly, it provides scalability in relation to the growth of data size, where Hadoop can scale to more computing nodes in the cluster to maintain the performance.
%
%Our framework benefits from using Hadoop in both data management and processing.
%First, since the input data is large it is desirable to store it on multiple machines.
%Second, by using Hadoop we can parallelize the computational tasks and make the data processing faster and more efficient.
%Further, by adopting the MapReduce computing paradigm, we can model each individual user's trajectory by treating user's unique ID as key, and summarize the movement flows among different spatial units by utilizing the ID of the spatial unit as key. The details will be introduced in the following section.

\section{Materials and Methods}
\subsection{Geo-located Twitter data}
Geo-located tweets are tweets appended with an additional geo-tag in the form of a pair of geographical coordinates, which represents the location a tweet was sent at. 
In this study, the geo-located tweets were downloaded using the Twitter Streaming API, where we specified a geographical bounding box as an area-of-interest to retrieve all the geo-located tweets that fall within it. 
To ensure complete coverage over the United States, we implemented a crawler that selects North America as the initial area-of-interest, where the geographical boundary is specified with lower left (latitude: 5.4, longitude: -167.3) and upper right (latitude: 83.2, longitude: -52.2).
The crawler is constantly running with over 2 million geo-located raw tweets ($\sim$2GB in size) collected per day.
We have collected more than 1.3 billion geo-located tweets from 1$^{st}$ January to 31$^{st}$ December, 2014 with 6,147,430 Twitter users and 1 TB in file size.

As a social media account is not equal to a real person in the physical world~\cite{tsou2015}, to ensure the data quality, the collected raw tweets were further filtered by the following steps: We first removed the duplicated messages\footnote{https://support.twitter.com/articles/18311-the-twitter-rules} in the dataset; and then we attempted to remove non-human users based on unusual relocating speed (Hawelka et al., 2014; Jurdak et al., 2015). 
In this case, we adopted the speed limit value as 240 m/s used in (Jurdak et al., 2015), where we examined all the consecutive locations of each user and excluded those with relocating speed over the limit.
Note that the original location information embedded in each geo-located tweet is given in units of latitude and longitude. To reduce the complexity in calculating distances and performing spatial aggregations directly based on geographical coordinates, we projected the all the points using North America Albers Equal Area Conic (ESRI:102008) coordinate system.
Finally, we used the geographic boundaries of the United States\footnote{https://www.census.gov/geo/maps-data/data/} (excluding Alaska and Hawaii) to further restrict the remaining tweets, where the technical details is presented in the following section. Based on these reinforcements, the dataset contains 1,052,861,000 tweets and 4,559,205 unique users.

\subsection{A scalable visual-analytics framework}
To address the data-intensive challenges, we have developed a scalable visual-analytics framework tailored to accommodate large volume of geo-located tweets for studying multi-level spatiotemporal Twitter user mobility patterns.
The scalable visual-analytics framework consists of two main units: (1) Data collection unit: a Twitter data crawler that continuously collects raw geo-located tweets using the Twitter Streaming API. (2) Data processing unit: a distributed computing environment using Apache Hadoop for modeling and extracting Twitter user movements, generating space-time user trajectories, and summarizing the movements at multiple spatiotemproal levels.
An overall system architecture of the framework that chains the three units is shown in Figure 1. The details regarding the function and implementation of each unit are presented in the next sections.

>>>>

%=================================================================
% Full title of the paper (Capitalized)
%\Title{Exploring multi-scale spatiotemporal Twitter user mobility patterns with a scalable visual-analytics framework}
\Title{Exploring Multi-Scale Spatiotemporal Twitter User Mobility Patterns with a Visual-Analytics Approach}

% Authors, for the paper (add full first names)
\Author{Junjun Yin $^{1}$ and Zhenhong Du $^{2,}$*}
% Authors, for metadata in PDF
\AuthorNames{Junjun Yin and Zhenhong Du}

@@

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{Understanding human mobility patterns is of great importance for urban planning, traffic management, and even marketing campaign.
However, the capability of capturing detailed human movements with fine-grained spatial and temporal granularity is still limited.
In this study, we extracted high-resolution mobility data from a collection of over 1.3 billion geo-located Twitter messages.
Regarding the concerns of infringement on individual privacy, such as the mobile phone call records with restricted access, the dataset is collected from publicly accessible Twitter data streams.
In this paper, we employed a visual-analytics approach to studying multi-scale spatiotemporal Twitter user mobility patterns in the contiguous United States during the year 2014.
Our approach included a scalable visual-analytics framework to deliver efficiency and scalability in filtering large volume of geo-located tweets, modeling and extracting Twitter user movements, generating space-time user trajectories, and summarizing multi-scale spatiotemporal user mobility patterns.
We performed a set of statistical analysis to understand Twitter user mobility patterns across multi-level spatial scales and temporal ranges.
In particular, Twitter user mobility patterns measured by the displacements and radius of gyrations of individuals revealed multi-scale or multi-modal Twitter user mobility patterns.
By further studying such mobility patterns in different temporal ranges, we identified both consistency and seasonal fluctuations regarding the distance decay effects in the corresponding mobility patterns.
At the same time, our approach provides a geo-visualization unit with an interactive 3D virtual globe web mapping interface for exploratory geo-visual analytics of the multi-level spatiotemporal Twitter user movements.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Understanding human mobility patterns is of great importance for a broad range of applications from urban planning~\cite{zheng2008understanding}, traffic management~\cite{jiang2009characterizing}, and even the spatial spread of epidemic diseases~\cite{belik2011natural}.
Earlier research efforts relied on low-resolution mobility data to understand human mobility patterns, such as using census records to study human migration patterns~\cite{greenwood1985human}, or delivering questionnaires and asking volunteers to report the track of bank notes to infer human travel patterns~\cite{brockmann2006scaling}.
However, such mobility data do not provide detailed human movements with fine-grained spatial and temporal granularity, which are usually aggregated and therefore are limited to capture mobility patterns of individuals~\cite{gonzalez2008understanding,Jurdak2015}.
In addition to the mobility data collected by GPS trackers~\cite{zheng2008understanding, rhee2011levy} and mobile phone call records~\cite{gonzalez2008understanding,sevtsuk2010does,kung2014exploring}, emerging as a new source for mobility data, today's pervasive Location Based Social Media (LBSM) platforms (e.g., Twitter and Foursquare) offer continuous spatial Big Data streams with massive amount of detailed and frequently updated user digital footprints in the form of real-world user trails and footprints~\cite{thatcher2014living}.
A significant advantage of utilizing LBSM data streams as proxies for studying human mobility patterns is the large spatial coverage.
For instance, researchers have used geo-located Twitter data for studying global mobility patterns~\cite{hawelka2014geo}, which is otherwise impossible for other mobility datasets (e.g., GPS traces and mobile phone call records).
Regarding the concerns of infringement on individual privacy, such as the mobile phone call records with restricted access~\cite{giannotti2008mobility,crampton2014collect,Jurdak2015}, the publicly available LBSM data streams offer unique opportunities for conducting reproducible and comparative scientific findings across different geographic regions.

Many recent studies have adopted the LBSM data streams to study human mobility patterns.
For example, they modeled and extracted trajectories of individuals and performed statistical analysis focusing on the distance decay effects in the collective user movements~\cite{gonzalez2008understanding}, which were used to reveal different travel modes~\cite{Jurdak2015}, travel demands~\cite{wu2014intra,hasan2013understanding}, and the impact of social connections~\cite{cho2011friendship}.
These studies have provided strong supports for using LBSM data as proxies for studying mobility patterns of individuals and valuable insights into human mobility dynamics.
However, the variations of movements in different spatial scales and temporal ranges are neglected in these studies, where the measurements of distances are either fixed in a certain time range or to a specific geographic region.
For instances, the examinations on whether there are temporal (e.g., monthly or seasonal) changes within the movements or how the observed mobility patterns vary across different spatial scales (e.g., intra- or inter city or national level), are lacking.
Such insights are critical to advance our understandings of the collective mobility patterns for various applications, such as examining the mobility patterns across different cities~\cite{noulas2012tale}, the spread patterns of disease~\cite{balcan2009multiscale, tamerius2011global} and touristic activities~\cite{hawelka2014geo}.
On the other hand, while the high-resolution spatiotemporal records from LBSM present unique research opportunities in this direction, the inherited large data volume poses significant data-intensive challenges for developing multi-scale spatiotemporal analysis approaches to dealing with the complexities in filtering movements of individuals, modeling and aggregating user trajectories at multiple spatial and temporal scales~\cite{tsou2015}.

In this paper, we have employed a visual-analytics approach to exploring the Twitter user mobility patterns across multi-level spatial scales and temporal ranges in the continuous United States (i.e., excluding Alaska and Hawaii) during the year 2014.
The mobility data is extracted from over 1.3 billion geo-located Twitter messages (i.e., tweets) from 1$^{st}$ January to 31$^{st}$ December, 2014 over North America with over 6 million Twitter users and over 1 TB in file size.
To address the data-intensive challenge embedded in this dataset, we have developed a scalable visual-analytics framework tailored to accommodate large volume of geo-located tweets.
This framework is implemented based on high-performance distributed computing environment using Apache Hadoop\footnote{http://hadoop.apache.org/}, which is an open source software framework to enable distributed processing of large datasets across computing clusters.
Enabled by this framework, we have performed a set of statistical analysis to understand multi-scale spatiotemporal Twitter user mobility patterns. 
We have modeled the frequency of Twitter users visiting different locations to study the collective user visiting behaviors, where we have identified temporal similarities in the distributions.
In particular, the Twitter user mobility patterns measured by the user displacements and radius of gyrations of individuals~\cite{gonzalez2008understanding} have revealed different groups of Twitter users with multi-scale or multi-modal mobility patterns and multiple travel modes~\cite{Jurdak2015}.
By further studying such mobility patterns in different temporal ranges, we have identified both consistency and seasonal fluctuations regarding the distance decay effects in the corresponding mobility patterns.
In particular, our approach provides an interactive 3D virtual globe web mapping interface to enable exploratory geo-visual analytics for understanding the detailed Twitter user movement flows within a given spatial scale and time window.

The remainder of this paper is organized as follows.
Section 2 describes the related work in the context of studying mobility patterns using LBSM data, in particular, the geo-located Twitter data.
We focus on research challenges in using visual-analytics methods to enable multi-scale spatiotemporal analysis with massive movement datasets, including data management, multi-level spatiotemporal user trajectory modeling and visualization.
Section 3 details the processes for extracting, aggregating and summarizing multi-level spatiotemporal Twitter user mobility patterns.
Section 4 presents the case study of performing visual-analytics for seeking multi-scale spatiotemporal Twitter mobility patterns in the continuous United States of year 2014.
Section 5 concludes the paper.

\section{Mobility patterns in Location Based Social Media data}
\subsection{Geo-located Twitter data for studying large-scale user movements}
To understand detailed mobility patterns of individuals, the capability of capturing human movements with fine-grained spatial and temporal granularity is critical.
In terms of collecting detailed mobility data of individuals, using GPS trackers tends to produce, to date, the most accurate records of individuals' movements regarding the accuracy of recorded user locations and update frequency~\cite{zheng2008understanding}.
However, the data are often limited in spatial scale (e.g. within a specific city or region) with a small group of people, for example, 226 and 182 volunteers participated in collecting such mobility data in~\cite{rhee2011levy} and~\cite{zheng2010geolife} respectively. Other than tracking people directly, the vehicle-based GPS traces are often tied to specific vehicles (e.g. taxi), which are only accessible for a certain group of people~\cite{kung2014exploring}.

Another approach from the literatures for studying human mobility is using mobile phone call data, such as Call Detail Records (CDR), where the locations of mobile users are estimated by cell tower triangulation with accuracy in the order of kilometers~\cite{gonzalez2008understanding,sevtsuk2010does,kung2014exploring}.
Such a dataset can cover relatively large spatial scale~\cite{becker2013human,sobolevsky2013delineating} (e.g., national level) and a large portion of the population in the study region~\cite{kung2014exploring}.
However, due to the concerns of infringement on individual privacy, mobile phone call data are not publicly accessible at all.
Even such data were obtained in the mentioned studies, they came from various service providers covering different groups of users.
These issues limit the capability for conducting reproducible scientific findings regarding mobility research, such as validating or extending the existing discoveries.

In this connection, it becomes increasingly popular for researchers to exploit the publicly accessible mobility data captured from today's pervasive Location Based Social Media (LBSM) platforms (e.g., Foursquare and Twitter).
LBSM enables users to attach their current location as a geo-tag to the message they post, which is derived from either the GPS or Wi-Fi positioning with a high position resolution down to 10 meters~\cite{Jurdak2015}.
A Big Data scenario emerges when millions social media users constantly posting messages.
In this study, geo-located Twitter data are chosen as a source for studying detailed mobility patterns.
Compared to other LBSM platforms, Twitter is one of the most popular platforms and is been actively used in many countries.
It provides a publicly accessible streaming API\footnote{https://dev.twitter.com/streaming/overview} for easy access to its data, in fact, many other LBSM data can be collected from the data streams, such as Foursquare check-in data~\cite{cranshaw2012livehoods,hasan2013understanding}.

However, it is worth noting that there are some limitations and complexities in directly using LBSM data for studying human mobility patterns.
For example, comparing to GPS traces, the update frequency of an individual's location varies depending on when a user is posting a new geo-located message or check-in at a new place.
There is a potential mismatch regarding the representativeness of the overall population as not all people use social media or send geo-located messages~\cite{kung2014exploring} and the demographic information of the Twitter users cannot be easily identified, the derived mobility patterns may lead to an over or under-representation of the real-world mobility patterns.
Many studies started to look into the demographic information of LBSM data, in particular Twitter data~\cite{mitchell2013geography,longley2015geotemporal}.
Although the used methods are still arguable, these issues certainly require us to pose stricter criteria in understanding human mobility patterns using geo-located Twitter data.
On the other hand, geo-located Twitter dataset presents some unique advantages that make it a valuable proxy for studying human mobility patterns.
For example, the high-resolution location information enables to identify multiple travel modes in user mobility patterns~\cite{Jurdak2015}; the large spatial coverage enables to study global mobility patterns~\cite{hawelka2014geo}, which is almost impossible for other mobility datasets.
More importantly, by continuously monitoring the geo-located Twitter data streams with large volumes of detailed and frequently updated spatiotemporal records of Twitter users, it offers a great deal of potential for studying mobility patterns of large groups of individuals at different spatial scales (e.g., movements across cities, states or even countries) and temporal gratuity (e.g., weekly, monthly, and seasonal movements), which is one of the motivations for this study.

\subsection{Data-intensive challenges for multi-scale geo-visual analytics}
Mobility data are essentially a collection of spatiotemporal records of people moving from one location to another across the geographic space.
To study mobility patterns of individuals, a space-time trajectory of each individual user should be modeled and constructed to quantify the collective movements over space and time.
Based on the extracted space-time trajectories, aforementioned studies are able to perform analysis, such as the measurements of user displacements and radius of gyrations of individuals, to study the mobility dynamics.
At the same time, space-time trajectory is one of the core concepts in H{\"a}gerstrand's time geography to understand the embedded spatiotemporal dynamics~\cite{hagerstrand1985time}, which has provided useful insights to explore movements across different geographical scales and temporal ranges.
For example, a geo-visualization approach was used to study human activity patterns, where user trajectories are mapped in a 3D space ordered by timestamps in the third dimension~\cite{kwan2004geovisualization}.
While such an approach enables visualization of individual trajectories, its capability is limited in dealing with large-volume movement datasets~\cite{andrienko2007designing}.
Instead of directly visualizing individual user trajectories, a space-time cube approach was proposed to analyzing and visualizing the collective trajectories.
It provides flexibilities in setting up both spatial scales and temporal ranges, and therefore is used to study mobility patterns across different spatial units (e.g. countries, states, and cities, etc.) and identify the changes over space and time~\cite{maceachren2001research, maceachren2004maps}.
In this regard, visual-analytics methods are proposed to help better convey the findings in terms of analyzing and visualizing multi-level spatiotemporal mobility patterns~\cite{andrienko2007designing,andrienko2007visual}.
Visual-analytics methods focus on the synergy of computational and analytical methods to reduce the visual clutter, where aggregation methods are suggested to perform grouping/dividing individual's moving trajectories at different spatial and temporal granularity, e.g., utilizing the space-time cube approach~\cite{andrienko2007designing}.
Employing visual-analytics methods dealing with massive movement datasets is not only beneficial for optimizing visualizations but also provides a great deal of flexibilities for performing statistical analysis in seeking mobility patterns with different level of spatiotemporal details.

However, in the context of studying mobility patterns using large volume of geo-located Twitter data, the inherited large data volume poses significant data intensive challenges for visual-analytics methods to scale with both the data volume and the computational requirements (e.g., movement extraction and trajectory modeling)~\cite{cao2014scalable}.
In particular, in our study, 1.3 billion geo-located tweets were collected with over 1 TB in file size.
To construct a space-time trajectory of an individual, it is necessary to go through the massive dataset to sort and update the trajectory whenever a new location is found.
Such a task is already computationally demanding, let along breaking the trajectories to construct space-time cube with multiple spatial scale and temporal ranges.
Indeed, developing a multi-scale spatiotemporal analysis approach is identified as one of the research challenges for dealing with social media Big Data~\cite{tsou2015}.
To address the data-intensive challenges, there is a need to develop a scalable visual-analytics framework tailored to accommodate large volume of geo-located tweets for studying multi-level spatiotemporal Twitter user mobility patterns.

\section{Materials and Methods}
\subsection{Geo-located Twitter data}
Geo-located tweets are tweets appended with an additional geo-tag in the form of a pair of geographical coordinates, which represents the location a tweet was sent at.
In this study, the geo-located tweets were downloaded using the Twitter Streaming API, where we specified a geographical bounding box as an area-of-interest to retrieve all the geo-located tweets that fall within it.
To ensure complete coverage over the continuous United States, we implemented a crawler that selects North America as the initial area-of-interest, where the geographical boundary is specified with lower left (latitude: 5.4, longitude: -167.3) and upper right (latitude: 83.2, longitude: -52.2).
The crawler is constantly running with over 2 million geo-located raw tweets ($\sim$2GB in size) collected per day.
We have collected more than 1.3 billion geo-located tweets from 1$^{st}$ January to 31$^{st}$ December, 2014 with 6,147,430 Twitter users and 1 TB in file size.

As a social media account is not equal to a real person in the physical world~\cite{tsou2015}, to ensure the data quality, the collected raw tweets were further filtered by the following steps: We first removed duplicated messages\footnote{https://support.twitter.com/articles/18311-the-twitter-rules} in the dataset; and then we removed non-human users based on the heuristic of unusual relocating speed discussed in~\cite{hawelka2014geo,Jurdak2015}. 
In this case, we adopted the speed limit value as 240 m/s used in~\cite{Jurdak2015}, where we examined all the consecutive locations of each user and excluded those with relocating speed over the limit.
Note that the original location information embedded in each geo-located tweet is given in units of latitude and longitude, the distance is calculated by the great-circle distance between two points on a sphere with the haversine formula.
Finally, we used the geographic boundaries of the continuous United States\footnote{https://www.census.gov/geo/maps-data/data/} (excluding Alaska and Hawaii) to further restrict the remaining tweets, where the technical details is presented in the following section. Based on these reinforcements, the dataset contains 1,052,861,000 tweets and 4,559,205 unique users.

\subsection{A scalable visual-analytics framework}
To address the data-intensive challenges, we have developed a scalable visual-analytics framework tailored to accommodate large volume of geo-located tweets for studying multi-level spatiotemporal Twitter user mobility patterns.
The scalable visual-analytics framework consists of two main units: (1) Data processing unit: a distributed computing environment using Apache Hadoop for modeling and extracting Twitter user movements, generating space-time user trajectories, and summarizing the movements at multiple spatiotemporal scales.
(2) Geo-visualization unit: an interactive 3D virtual globe web mapping interface for exploratory geo-visual analytics for understanding the detailed Twitter user movement flows across different spatial scales and temporal ranges.

Apache Hadoop combines a distributed file system, namely Hadoop Distributed File System(HDFS)~\cite{shvachko2010hadoop} with MapReduce programming paradigm~\cite{dean2008mapreduce}, which can be applied to a wide range of data-intensive problems.
Our framework benefits from using Hadoop in both data management and processing.
First, since the input data is large it is desirable to store it on multiple machines.
This provides scalability in relation to the growth of data size, where Hadoop can scale to more computing nodes in a cluster to maintain the performance.
Second, by using Hadoop we can parallelize the computational tasks, where MapReduce breaks the entire computation into small tasks and schedule them among different computing nodes, to make the data processing faster and more efficient.
An overall system architecture of the framework is shown in Figure~\ref{fig:overall_archi}.
The details regarding the function and implementation of each unit are presented in the next sections.

\begin{figure}[ht]
\centering
\includegraphics[width=0.6\linewidth]{./figures/Overall_Architecture222}
\caption{The overall system architecture of the framework}
\label{fig:overall_archi}
\end{figure}
\FloatBarrier
